Differential cultural reproduction of skill level and mental templates in Late Acheulean handaxe morphology: Archaeological and experimental insights

### 2023-01-07 manuscript/manuscript.Rmd
added a few more references.

The morphological variability of Acheulean handaxes has been one of the most well-studied and well-published topics in paleolithic archaeology [@white1998; @earlyhu1998; @key2019]. Despite the recurrent narrative emphasizing the homogeneity and longevity of handaxe assemblages on a global scale and the conservatism behind this phenomenon that evokes genetic explanations [@corbey2016; @corbey2020; @richerson2005; @sterelny2004], many researchers have recognized the diversity within what has been deemed as a unified Acheulean "tradition" and tried to dissect the sources and meaning of this variation [@lycett2008; @nowell2010; @nowell2002; @sharon2011]. More specifically, a complex suite of interconnecting factors [@lycett2015] have been identified to contribute to handaxe morphological variation, including but not limited to raw material variability [@sharon2008; @eren2014; @lycett2016; @mcnabb2015], percussor properties [@shipton2009], functional differences [@machin2007; @key2016; @white2018; @key2017; @lycett2008], reduction method/intensity [@shipton2009; @shipton2015], time budgets [@schillingerConsideringRoleTime2014], learning processes [@lycett2016; @kempe2012], social signaling [@kohn1999; @spikins2012], aesthetic preferences [@gowlett2021; @letensorer2006], knapper skill levels [@stout2014; @herzlinger2017; @caruana2021], and mental templates [@garcía-medrano2019; @hutchence2021; @schillinger2017]. From this extensive list, knapper skill levels and mental templates have been repeatedly mentioned and discussed in the now extensive corpus of handaxe studies, and Boxgrove handaxes have been one of the most studied assemblages from these two angles. Of particular attention here are the experimental works conducted by Stout et al. [-@stout2014] focusing on inferring knapping skill level and Garcia-Medrano et al. [-@garcía-medrano2019] identifying the mental template of the Boxgrove assemblage. Our paper combines these two perspectives and provides novel insights to the same archaeological assemblage by comparing it with experimentally made handaxes.

>>>>

The morphological variability of Acheulean handaxes has been one of the most well-studied and well-published topics in paleolithic archaeology [@white1998; @earlyhu1998; @key2019]. Despite the recurrent narrative emphasizing the homogeneity and longevity of handaxe assemblages on a global scale and the conservatism behind this phenomenon that evokes genetic explanations [@corbey2016; @corbey2020; @richerson2005; @sterelny2004], many researchers have recognized the diversity within what has been deemed as a unified Acheulean "tradition" and tried to dissect the sources and meaning of this variation [@lycett2008; @nowell2010; @nowell2002; @sharon2011; @moncel2018a; @moncel2018b; @moncel2018c]. More specifically, a complex suite of interconnecting factors [@lycett2015] have been identified to contribute to handaxe morphological variation, including but not limited to raw material variability [@sharon2008; @eren2014; @lycett2016; @mcnabb2015], percussor properties [@shipton2009], functional differences [@machin2007; @key2016; @white2018; @key2017; @lycett2008], reduction method/intensity [@shipton2009; @shipton2015], time budgets [@schillingerConsideringRoleTime2014], learning processes [@lycett2016; @kempe2012], social signaling [@kohn1999; @spikins2012], aesthetic preferences [@gowlett2021; @letensorer2006], knapper skill levels [@stout2014; @herzlinger2017; @caruana2021], and mental templates [@garcía-medrano2019; @hutchence2021; @schillinger2017]. From this extensive list, knapper skill levels and mental templates have been repeatedly mentioned and discussed in the now extensive corpus of handaxe studies, and Boxgrove handaxes have been one of the most studied assemblages from these two angles. Of particular attention here are the experimental works conducted by Stout et al. [-@stout2014] focusing on inferring knapping skill level and Garcia-Medrano et al. [-@garcía-medrano2019] identifying the mental template of the Boxgrove assemblage. Our paper combines these two perspectives and provides novel insights to the same archaeological assemblage by comparing it with experimentally made handaxes.
